Dominance Regions for Imaginary $\bfg$-vectors

### 2021-06-28 dominance_regions.tex
Added some results

\newcommand{\bfc}{\mathbf{c}}
\newcommand{\bfg}{\mathbf{g}}

@@

@@

  \section{Lattice Mutations}
  Let $B=(b_{ij})$ be an $n\times n$ skew-symmetrizable integer matrix, i.e. there exists a diagonal integer matrix $D=\diag(d_1,\ldots,d_n)$ so that $DB$ is skew-symmetric, more precisely we have $d_i b_{ij}=-d_j b_{ji}$.
  For $b\in\ZZ$, write $[b]_+=max(b,0)$.
  Given a sign $\varepsilon\in\{\pm\}$ and $k\in\{1,\ldots,n\}$, define a matrix $E_{k,\varepsilon}=(e_{ij})$ with
  \begin{equation}
    \label{eq:left mutation matrix}
    e_{ij}=\begin{cases} 1 & \text{if $i=j\ne k$;}\\ -1 & \text{if $i=j=k$;}\\ [\varepsilon b_{ik}]_+ & \text{if $i\ne j=k$;}\\ 0 & \text{otherwise;} \end{cases}
  \end{equation}
  and a matrix $F_{k,\varepsilon}=(f_{ij})$ with
  \begin{equation}
    \label{eq:right mutation matrix}
    f_{ij}=\begin{cases} 1 & \text{if $k\ne i=j$;}\\ -1 & \text{if $k=i=j$;}\\ [-\varepsilon b_{kj}]_+ & \text{if $k=i\ne j$;}\\ 0 & \text{otherwise.} \end{cases}
  \end{equation}
  Observe that $E^2_{k,\varepsilon}=\Id$ and $F^2_{k,\varepsilon}=\Id$ for any $k$ and any choice of $\varepsilon$.

@@

  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}
  For $\lambda\in\RR^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\RR^n$, define the \emph{dominance region} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.
  \begin{lemma}
    The sets $\cP(\lambda)$ are polytopes.
  \end{lemma}

@@

>>>>

\newcommand{\bfc}{\boldsymbol{c}}
\newcommand{\bfg}{\boldsymbol{g}}
\newcommand{\bfr}{\boldsymbol{r}}
\newcommand{\bfx}{\boldsymbol{x}}
\newcommand{\bfy}{\boldsymbol{y}}

@@

  \section{Questions}
  \begin{enumerate}
    \item To what extent can we explicitly/combinatorially describe the eigenvalues/eigenvectors of the twist automorphism/Auslander-Reiten translation acting on imaginary $\bfg$-vectors given a maximal green sequence? 
    \item What role do coefficients play in the dominance order on pointed bases?
    \item Is any change required in the theory to handle quantum/generalized or both cases together?
    \item What do our results mean geometrically in terms of the cluster variety?  Is this constructing an affine chart?
  \end{enumerate}

@@

  \section{Lattice Mutations and Linear Inequalities}
  Let $B=(b_{ij})$ be an $n\times n$ skew-symmetrizable integer matrix, i.e. there exists a diagonal integer matrix $D=\diag(d_1,\ldots,d_n)$ so that $DB$ is skew-symmetric, more precisely we have $d_i b_{ij}=-d_j b_{ji}$.
  Write $\Im(B)$ for the span of the columns of $B$ inside $\RR^n$ and $\Im^+(B)$ for their non-negative span.
  Define $\Im_\ZZ(B)$ and $\Im^+_\ZZ(B)$ analogously over the integers.
  The matrix $B$ is \emph{acyclic} if there is no sequence $i_1,\ldots,i_m,i_{m+1}=i_1$ so that $b_{i_\ell i_{\ell+1}}>0$ for $1\le\ell\le m$.
  In the case when $B$ is acyclic, there exists a labeling $\{k_1,\ldots,k_n\}=\{1,\ldots,n\}$ so that $\ell<\ell'$ implies $b_{k_\ell k_{\ell'}}\ge 0$, such a sequence is called \emph{sink-adapted}.

  For $b\in\ZZ$, write $[b]_+=max(b,0)$.
  Given a sign $\varepsilon\in\{\pm\}$ and $k\in\{1,\ldots,n\}$, define a matrix $E_{k,\varepsilon}=(e_{ij})$ with
  \begin{equation}
    \label{eq:left mutation matrix}
    e_{ij}=\begin{cases} 1 & \text{if $i=j\ne k$;}\\ -1 & \text{if $i=j=k$;}\\ [\varepsilon b_{ik}]_+ & \text{if $i\ne j=k$;}\\ 0 & \text{otherwise;} \end{cases}
  \end{equation}
  and a matrix $F_{k,\varepsilon}=(f_{ij})$ with
  \begin{equation}
    \label{eq:right mutation matrix}
    f_{ij}=\begin{cases} 1 & \text{if $k\ne i=j$;}\\ -1 & \text{if $k=i=j$;}\\ [-\varepsilon b_{kj}]_+ & \text{if $k=i\ne j$;}\\ 0 & \text{otherwise.} \end{cases}
  \end{equation}
  Observe that $E^2_{k,\varepsilon}=\Id$ and $F^2_{k,\varepsilon}=\Id$ for any $k$ and any choice of $\varepsilon$.

@@

  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}
  \begin{lemma}
    Suppose $B$ is acyclic with sink-adapted sequence $k_1,\ldots,k_n$.
    Consider $t_1,\ldots,t_n\in\TT_n$ so that $t_{\ell-1}$ is joined to $t_\ell$ by an edge in $\TT_n$ labeled by $k_\ell$ for $1\le\ell\le n$.
    For a sequence of signs $\varepsilon_1,\ldots,\varepsilon_n$, the product $E^{t_{n-1}}_{k_n,\varepsilon_n}\cdots E^{t_0}_{k_1,\varepsilon_1}$ is given by $A=(a_{ij})$ where
    \[a_{k_{\ell'} k_\ell}=\sum_{U=\{\ell=\ell_1<\ldots<\ell_t\}\subseteq[\![\ell,n]}\prod_{s=1}^t \begin{cases} [-\varepsilon_{\ell_t}b_{k_{\ell'} k_{\ell_t}}]_+ & \text{if $s=t$ and $\ell_t>\ell'$}\\ -1 & \text{if $s=t$ and $\ell_t=\ell'$}\\ [\varepsilon_{\ell_s} b_{k_{\ell_{s+1}} k_{\ell_s}}]_+ & \text{if $s<t$ and $\ell_t\ge\ell'$}\\ 0 & \text{if $\ell'>\ell_t$} \end{cases} \]

    Then $\phi_{t_n}$ is the following piecewise-linear automorphism:
    \begin{equation}
      \label{eq:tropical twist}
      \text{(This is going to require casework with the possible signs, we'll really only need it for imaginary $\bfg$-vectors.)}
    \end{equation}
  \end{lemma}
  \begin{proof}
    We prove the following modified formula for the product $E^{t_{m-1}}_{k_m,\varepsilon_n}\cdots E^{t_0}_{k_1,\varepsilon_1}$ with $1\le m\le n$ by induction on $m$.
    Writing this matrix as $A^{(m)}=(a^{(m)}_{ij})$, we claim that
    \[a^{(m)}_{k_{\ell'} k_\ell}=\sum_{U=\{\ell=\ell_1<\ldots<\ell_t\}\subseteq[\![\ell,n]}\prod_{s=1}^t \begin{cases} [-\varepsilon_{\ell_t}b_{k_{\ell'} k_{\ell_t}}]_+ & \text{if $s=t$ and $m\ge\ell_t>\ell'$}\\ -1 & \text{if $s=t$ and $m\ge\ell_t=\ell'$}\\ [\varepsilon_{\ell_s} b_{k_{\ell_{s+1}} k_{\ell_s}}]_+ & \text{if $s<t$ and $\ell_t\ge\ell'$}\\ [\varepsilon_\ell b_{k_{\ell'} k_\ell}]_+ & \text{if $\ell'>m\ge\ell_t=\ell$}\\ 1 & \text{if $\ell=\ell_t=\ell'>m$}\\ 0 & \text{if $m\ge\ell'>\ell_t$ or $\ell_t>\ell'>m$} \end{cases} \]
    Observe that the additional conditions vanish when $m=n$, proving the desired result once this formula is established.

    Since the sequence $k_1,\ldots,k_n$ is sink-adapted, the matrix $E^{t_{\ell-1}}_{k_\ell,\varepsilon_\ell}$, for $1\le\ell\le n$, has entries $e^{(\ell)}_{ij}$ given by
    \begin{equation}
      \label{eq:sink mutation matrix}
      e^{(\ell)}_{ij}=\begin{cases} 1 & \text{if $i=j\ne k_\ell$;}\\ -1 & \text{if $i=j=k_\ell$;}\\ [-\varepsilon_\ell b_{ik_\ell}]_+ & \text{if $i=k_{\ell'}$, $j=k_\ell$, $\ell>\ell'$;}\\ [\varepsilon_\ell b_{ik_\ell}]_+ & \text{if $i=k_{\ell'}$, $j=k_\ell$, $\ell'>\ell$;}\\ 0 & \text{otherwise;} \end{cases}
    \end{equation}
    the entries for $\ell'<\ell$ picking up this sign from the prior matrix mutations.
    From this, it is easy to see that $a^{(1)}_{k_{\ell'} k_\ell}=e^{(1)}_{k_{\ell'} k_\ell}$ for $\ell,\ell'\in[1,n]$, this establishes the base of the induction on $m$.

    Now suppose the matrix $A^{(m)}$ is given as above.
    We need to check that 
    \begin{equation}
      \label{eq:twist matrix recursion}
      a^{(m+1)}_{k_{\ell'} k_\ell}=\sum\limits_{\ell''=1}^n e^{(m+1)}_{k_{\ell'} k_{\ell''}} a^{(m)}_{k_{\ell''} k_\ell}.
    \end{equation}
    It will be convenient to distinguish cases:
    \begin{itemize}
      \item For $\ell'=m+1$, the right hand side of \eqref{eq:twist matrix recursion} becomes $-a^{(m)}_{k_{m+1} k_\ell}$.
      \item If $\ell'>m+1$, the right hand side of \eqref{eq:twist matrix recursion} becomes $a^{(m)}_{k_{\ell'} k_\ell}+[\varepsilon_{m+1} b_{k_{\ell'} k_{m+1}}]_+ a^{(m)}_{k_{m+1} k_\ell}$.
      \item If $m+1>\ell'$, the right hand side of \eqref{eq:twist matrix recursion} becomes $a^{(m)}_{k_{\ell'} k_\ell}+[-\varepsilon_{m+1} b_{k_{\ell'} k_{m+1}}]_+ a^{(m)}_{k_{m+1} k_\ell}$.
    \end{itemize}
  \end{proof}

  For $\lambda\in\RR^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\RR^n$, define the \emph{dominance region} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.
  \begin{lemma}
    The sets $\cP(\lambda)$ are polytopes.
  \end{lemma}

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A$ be an invertible $n\times n$ matrix.
    Under the left action of $A$, say $A\bfx=\bfy$ for $\bfx,\bfy\in\RR^n$, the region inside $\RR^n$ defined by the inequality $\bfr\bullet(\bfx-\lambda)\le t$ with $\lambda,\bfr\in\RR^n$ and $t\in\RR$ is transformed into the region defined by the inequality \[(A^{-T}\bfr)\bullet(\bfy-A\lambda)\le t.\]
  \end{lemma}
  \begin{proof}
    This is immediate from the equalities
    \[\bfr\bullet(\bfx-\lambda)=\bfr^T(\bfx-\lambda)=\bfr^TA^{-1}A(\bfx-\lambda)=(A^{-T}\bfr)^T(A\bfx-A\lambda)=(A^{-T}\bfr)\bullet(\bfy-A\lambda).\]
  \end{proof}

@@

  \begin{lemma}
    One column is always in the positive span of the other columns.
  \end{lemma}
